---
title: "Using Socio-Demographic Data in Practice"
author: "Marga Torre"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
---

<style>
body {
text-align: justify}
</style> 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, message=FALSE, warning=FALSE, knitr.purl.inline = TRUE)
```

```{r include=FALSE}
library(haven)
library(tidyverse)
library(knitr)
install.packages("essurvey")
library(gmodels)
library(ggplot2)
library(openxlsx)
library(skimr)
library(ggplot2)
library(janitor)
library(readr)
library(dplyr)
library(labelled)
library(essurvey)
```

## European Social Survey
The European Social Survey (ESS) is an academically driven cross-national survey that has been conducted across Europe since its establishment in 2001. Every two years, face-to-face interviews are conducted with newly selected, cross-sectional samples. (Taken from the [ESS website](http://www.europeansocialsurvey.org/about-ess))


## Read data
```{r}
library(readr)
data_full <- read_csv("ESS_es.csv")
dim(data_full)
```

You can use the [`essurvey` package](https://cran.r-project.org/web/packages/essurvey/index.html) to retrieve the data from the website. However, since it is not working properly at the moment while I'm preparing for the class, I have downloaded and imported the data in the conventional way.


### Select variables
ESS round + ID + socio-demographic profile, including: gender, age of the respondent, years of education, level of education, occupation, household's total net income and citizenship. Please be aware that __some variable names may change from one round to another__.

```{r}
data = data_full |> select(essround, idno, gndr, agea, eduyrs, eisced, iscoco, isco08, hinctnt, hinctnta)
head(data)
```

agea: age of respondents
edlves: highest level of educaiton, Spain
iscoco (w1, w5), isco08 (w6 - w10): occupation
hinctnt, hinctnta: household income


## Data cleaning and data wrangling 
### Gender

Generate a new variable named `men` and assign it the following values. Label the categories of the variable  and check the results using a table.

-   1=man
-   0=woman
-   other=NA

```{r}
data_I = data
data_I$men = if_else(data$gndr == 1, "men", 
                      if_else(data$gndr == 2, "women", NA))

table(data$essround, data_I$men)


  
data = data |> 
  mutate(male = car::recode(gndr, "2=0; 9=NA"),
         male = factor(male, levels = c(0,1), labels = c("female", "male")))
table(data$male)
table(data$essround, data$male)

data = data |> relocate(male, .before=gndr)
```

### Age

Generate a continuous variable named `age`, with the maximum value being 103 years:

```{r}
data <- data |> mutate(age = replace(agea, agea > 103, NA))

summary(data$age)
```

Using the continuous variable `age`, generate a new variable named `age_rec` with the following categories (levels). Don't forget to label the categories of the variable and check the results using a table.

1.  "<20"
2.  "20-29"
3.  "30-44"
4.  "45-64"
5.  ">=65" 

```{r}
data <- data |> 
  mutate(age_rec = cut(agea, breaks = c(0, 19, 29, 44, 64, 103),
                       labels = c("<20","20-29","30-44","45-64",">=65")))

data = 

with(data, table(essround, age_rec))
```

### Education
Generate a variable named `years_schooling` defined as a continuous variable. 

```{r}
data = data |> mutate(years_schooling = if_else(data$eduyrs < 77, eduyrs, NA))



head(data)
```

Based on the _EISCED_ variable, generate a variable named _educ_level_ with the below categories. Ensure that the categories of the new variable are labeled  and check the results using a table.

1.  Low education
2.  Medium education
3.  High education

```{r}
data = data |> mutate(
  educ_level = case_when(
    eisced %in% c(1,2) ~ "Low education",
    eisced %in% c(3,4) ~ "Medium education",
    eisced %in% c()
  )
)
  
```


### Household's total net income
Generate a scale variable named hincome displaying income deciles. Note that the variable name changes from wave 3 to wave 4.

```{r}
data = data |> mutate(hincome = case_when(
  essround <= 3 ~ hinctnt,
  essround >= 4 ~ hinctnta,
  TRUE ~ NA_real_
))
data = data |> relocate(hincome, .before = hinctnt)
with(data, table(essround, hincome))
```


### Occupational categories
Using the variables _iscoco_ (waves <=5) and _isco08_ (waves >5), generate a new variable named `occup_cat` with the following levels:

1.    Managers
2.    Professionals
3.    Technicians
4.    Clerical
5.    Service
6.    Agricultural
7.    Blue-collar


```{r}

```


## OPTIONAL: Social class - Daniel Oesch 

Replicate Daniel Oesch's files __to construct the 5-class schemes__ for the interviewee's class position. Please note that starting from wave 6, there has been a change in the occupation classification from `iscoco` to `isco08`. Use [the scripts provided by Daniel Oesch](https://people.unil.ch/danieloesch/scripts/)

References:

-   Oesch, D. (2006a) "Coming to grips with a changing class structure" _International Sociology_ 21 (2): 263-288.
-   Oesch, D. (2006b) "Redrawing the Class Map. Stratification and Institutions in Britain, Germany, Sweden and Switzerland", Basingstoke: Palgrave Macmillan.


#### 5-Class schema constructed

1.    Higher-grade service class
2.    Lower-grade service class
3.    Small business owners
4.    Skilled workers
5.    Unskilled workers

