---
title: "Using Socio-Demographic Data in Practice"
author: "Marga Torre"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
---

<style>
body {
text-align: justify}
</style> 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, message=FALSE, warning=FALSE, knitr.purl.inline = TRUE)
```

```{r include=FALSE}
library(haven)
library(tidyverse)
library(knitr)
library(gmodels)
library(ggplot2)
library(openxlsx)
library(skimr)
library(ggplot2)
library(janitor)
library(readr)
library(dplyr)
library(labelled)
library(essurvey)
```

## European Social Survey
The European Social Survey (ESS) is an academically driven cross-national survey that has been conducted across Europe since its establishment in 2001. Every two years, face-to-face interviews are conducted with newly selected, cross-sectional samples. (Taken from the [ESS website](http://www.europeansocialsurvey.org/about-ess)). You can access the questionnarie from [here](https://www.europeansocialsurvey.org/methodology/ess-methodology/source-questionnaire).


## Read data
```{r}
data_full <- read_csv("ESS_es.csv")
dim(data_full)

```

You can use the [`essurvey` package](https://cran.r-project.org/web/packages/essurvey/index.html) to retrieve the data from the website. However, since it is not working properly at the moment while I'm preparing for the class, I have downloaded and imported the data in the conventional way.


### Select variables
ESS round + ID + socio-demographic profile, including: gender, age of the respondent, years of education, level of education, occupation, household's total net income and citizenship. Please be aware that __some variable names may change from one round to another__.

```{r}
data<- data_full |> select (essround, idno, agea, eduyrs, eisced, gndr, iscoco, isco08, hinctnt, hinctnta)
```

## Data cleaning and data wrangling 
### Gender

Generate a new variable named `men` and assign it the following values. Label the categories of the variable  and check the results using a table.

-   1=man
-   0=woman
-   other=NA

```{r}
data <- data |>
  mutate(
    male = car::recode(gndr, "2=0; 9=NA"),
    male = factor(male, levels = c(0, 1), labels = c("female", "male"))
  )

table(data$essround, data$male)
```

### Age

Generate a continuous variable named `age`, with the maximum value being 103 years:

```{r}
data <- data |>
  mutate(
    age = replace(agea, agea > 103, NA)
  )

summary(data$age) # 999 and other values above 
```

Using the continuous variable `age`, generate a new variable named `age_rec` with the following categories (levels). Don't forget to label the categories of the variable and check the results using a table.

1.  "<20"
2.  "20-29"
3.  "30-44"
4.  "45-64"
5.  ">=65" 

```{r}
data <- data |>
  mutate(
    age_rec = cut(agea, breaks = c(0, 19, 29, 44, 64, 103),
                  labels = c("<20", "20-29", "30-44", "45-64", ">=65"))
  )

with(data, table(essround, age_rec))
```

### Education

Generate a variable named `years_schooling` defined as a continuous variable. 

```{r}
data <- data |>
  mutate(
    years_schooling = replace(eduyrs, eduyrs > 60, NA)
  )

summary(data$years_schooling)
```

Based on the _EISCED_ variable, generate a variable named _educ_level_ with the below categories. Ensure that the categories of the new variable are labeled  and check the results using a table.

1.  Low education
2.  Medium education
3.  High education

```{r}
data <- data |>
  mutate(
    educ_level = case_when(
      eisced %in% c(1, 2) ~ "1",
      eisced %in% c(3, 4) ~ "2",
      eisced %in% c(5, 6, 7) ~ "3",
      eisced %in% c(0, 55, 77, 88, 99) ~ NA_character_,
      TRUE ~ NA_character_
    )
  ) |>
  mutate(
    educ_level = fct_recode(educ_level,
                            "Low education" = "1",
                            "Secondary education" = "2",
                            "High education" = "3")
  )

with(data, table(essround, educ_level))
```


### Household's total net income
Generate a scale variable named hincome displaying income deciles. Note that the variable name changes from wave 3 to wave 4.

First, we harmonize the name of the variable
```{r}
data <- data |>
  mutate(hincome = case_when(
    essround <= 3 ~ hinctnt,
    essround >= 4 ~ hinctnta,
    TRUE ~ NA_real_
  )) 
 # |> select(-hinctnt, -hinctnta) # if you want to remove previous vars
```

Next, we clean the variable `hincome` by removing non-valid answers.
```{r}
data <- data |>
  mutate(hincome = ifelse(hincome > 10, NA, hincome))

with(data, table(essround, hincome))
```


### Occupational categories

Using the variables _iscoco_ (waves <=5) and _isco08_ (waves >5), generate a new variable named `occup_cat` with the following levels:

1.    Managers
2.    Professionals
3.    Technicians
4.    Clerical
5.    Service
6.    Agricultural
7.    Blue-collar

As in the previous case, we first harmonize the name of the variable:

```{r}
data <- data |>
  mutate(occup_cat = case_when(
    essround <= 5 ~ iscoco,
    essround >= 6 ~ isco08,
    TRUE ~ NA_real_
  )) 
 # |> select(-iscoco, -ISCO08) # if you want to remove previous vars
```

Next, we remove non-valid entries, define the levels, and label them.
```{r}
data <- data |> 
  mutate(
    occup_cat = replace(occup_cat, which(occup_cat > 10000), NA),
    occup_cat = cut(
      occup_cat,
      breaks = c(0, 1999, 2999, 3999, 4999, 5999, 6999, 9999),
      labels = c(
        "managers",
        "professionals",
        "technicians",
        "clerical",
        "service",
        "agricultural",
        "blue-collar")))
```


Let's check the new variable by round:
```{r}
with(data, table(essround, occup_cat))
```

## Social class - Daniel Oesch

We are going to replicate Daniel Oesch's files __to construct the 5-class schemes__ for the interviewee's class position. Please note that starting from wave 6, there has been a change in the occupation classification from `iscoco` to `isco08`. This process has been carried out using [the scripts provided by Daniel Oesch](https://people.unil.ch/danieloesch/scripts/)

References:

-   Oesch, D. (2006a) "Coming to grips with a changing class structure" _International Sociology_ 21 (2): 263-288.
-   Oesch, D. (2006b) "Redrawing the Class Map. Stratification and Institutions in Britain, Germany, Sweden and Switzerland", Basingstoke: Palgrave Macmillan.


#### 16-Class schema constructed

1.    Large employers
2.    Self-employed professionals
3.    Small business owners with employees
4.    Small business owners without employees
5.    Technical experts
6.    Technicians
7.    Skilled manual
8.    Low-skilled manual
9.    Higher-grade managers and administrators
10.   Lower-grade managers and administrators
11.   Skilled clerks
12.   Unskilled clerks
13.   Socio-cultural professionals
14.   Socio-cultural semi-professionals
15.   Skilled service
16.   Low-skilled service

#### 8-Class schema constructed

1.    Self-employed professionals and large employers
2.    Small business owners
3.    Technical (semi-)professionals
4.    Production workers
5.    (Associate) managers
6.    Clerks
7.    Socio-cultural (semi-)professionals
8.    Service workers

#### 5-Class schema constructed

1.    Higher-grade service class
2.    Lower-grade service class
3.    Small business owners
4.    Skilled workers
5.    Unskilled workers


### Respondent's Oesch class position
Variables used to construct class variable for respondents: __iscoco, isco08, emplrel, emplno__. Let's add the new variables variable to our data set. To avoid losing all the variables we have created so far, we add variables to the initial selection.


```{r}
data2<-data_full |> select (idno, essround, emplrel, emplno) 
data <-left_join(data, data2, by=c("essround", "idno"))

```

### Waves 1-5
Let's start with waves 1-5. Recode occupation variable using _iscoco_ (isco88 4-digit) for respondents
```{r}
data$isco_mainjob <- data$iscoco
data$isco_mainjob[is.na(data$isco_mainjob)] <- -9
var_label(data$isco_mainjob) <- "Current occupation of respondent - isco88 4-digit"
```


Recode employment status for respondents. First, we create the variable _emplrel_r1_
```{r}
data$emplrel_r1 <- data$emplrel
data$emplrel_r1[is.na(data$emplrel_r1)] <- 9
val_label(data$emplrel_r1, 9) <- "Missing"
table(data$emplrel_r1)

```

Second, we create the variable _emplno_r1_
```{r}
data$emplno_r1 <- data$emplno
data$emplno_r1[is.na(data$emplno_r1)] <- 0
data$emplno_r1[data$emplno_r1 >= 1 & data$emplno_r1 <= 9] <- 1
data$emplno_r1[data$emplno_r1 >= 10 & data$emplno_r1 <= 66665] <- 2
val_labels(data$emplno_r1) <- c("0 employees" = 0,
                           "1-9 employees" = 1,
                           "10+ employees" = 2)
table(data$emplno_r1)

```

Finally, by combining the last two, we create the variable _selfem_mainjob_
```{r}
data$selfem_mainjob <- NA
data$selfem_mainjob[data$emplrel_r1 == 1 | data$emplrel_r1 == 9] <- 1
data$selfem_mainjob[data$emplrel_r1 == 2 & data$emplno_r1 == 0] <- 2
data$selfem_mainjob[data$emplrel_r1 == 3] <- 2
data$selfem_mainjob[data$emplrel_r1 == 2 & data$emplno_r1 == 1] <- 3
data$selfem_mainjob[data$emplrel_r1 == 2 & data$emplno_r1 == 2] <- 4
val_labels(data$selfem_mainjob) <- c("Not self-employed" = 1,
                                  "Self-empl without employees" = 2,
                                  "Self-empl with 1-9 employees" = 3,
                                  "Self-empl with 10 or more" = 4)
var_label(data$selfem_mainjob) <- "Employment status for respondants"
table(data$selfem_mainjob)
```



Create Oesch class schema for respondents
We start creating the variable _class16_r1_
```{r}
data$class16_r1 <- -9

```

Large employers (1)
```{r}
data$class16_r1[data$selfem_mainjob == 4] <- 1

```

Self-employed professionals (2)

```{r}
data$class16_r1[(data$selfem_mainjob == 2 | data$selfem_mainjob == 3) & data$isco_mainjob >= 2000 & data$isco_mainjob <= 2229] <- 2
data$class16_r1[(data$selfem_mainjob == 2 | data$selfem_mainjob == 3) & data$isco_mainjob >= 2300 & data$isco_mainjob <= 2470] <- 2
```


Small business owners with employees (3)
```{r}
data$class16_r1[data$selfem_mainjob == 3 & data$isco_mainjob >= 1000 & data$isco_mainjob <= 1999] <- 3
data$class16_r1[data$selfem_mainjob == 3 & data$isco_mainjob >= 3000 & data$isco_mainjob <= 9333] <- 3
data$class16_r1[data$selfem_mainjob == 3 & data$isco_mainjob == 2230] <- 3

```


Small business owners without employees (4)

```{r}
data$class16_r1[data$selfem_mainjob == 2 & data$isco_mainjob >= 1000 & data$isco_mainjob <= 1999] <- 4
data$class16_r1[data$selfem_mainjob == 2 & data$isco_mainjob >= 3000 & data$isco_mainjob <= 9333] <- 4
data$class16_r1[data$selfem_mainjob == 2 & data$isco_mainjob == 2230] <- 4
```

Technical experts (5)

```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2100 &  data$isco_mainjob <= 2213] <- 5
```

Technicians (6)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3100 &  data$isco_mainjob <= 3152] <- 6
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3210 &  data$isco_mainjob <= 3213] <- 6
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 3434] <- 6
```


Skilled manual (7)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 6000 &  data$isco_mainjob <= 7442] <- 7
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 8310 &  data$isco_mainjob <= 8312] <- 7
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 8324 &  data$isco_mainjob <= 8330] <- 7
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 8332 &  data$isco_mainjob <= 8340] <- 7
```


Low-skilled manual (8)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 8000 &  data$isco_mainjob <= 8300] <- 8
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 8320 &  data$isco_mainjob <= 8321] <- 8
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 8331] <- 8
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 9153 &  data$isco_mainjob <= 9333] <- 8
```


Higher-grade managers and administrators (9)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 1000 &  data$isco_mainjob <= 1239] <- 9
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2400 &  data$isco_mainjob <= 2429] <- 9
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2441] <- 9
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2470] <- 9
```


Lower-grade managers and administrators (10)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 1300 &  data$isco_mainjob <= 1319] <- 10
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3400 &  data$isco_mainjob <= 3433] <- 10
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3440 &  data$isco_mainjob <= 3450] <- 10
```


Skilled clerks (11)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 4000 &  data$isco_mainjob <= 4112] <- 11
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 4114 &  data$isco_mainjob <= 4210] <- 11
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 4212 &  data$isco_mainjob <= 4222] <- 11
```

Unskilled clerks (12)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 4113] <- 12
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 4211] <- 12
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 4223] <- 12
```

Socio-cultural professionals (13)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2220 &  data$isco_mainjob <= 2229] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2300 &  data$isco_mainjob <= 2320] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2340 &  data$isco_mainjob <= 2359] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2430 &  data$isco_mainjob <= 2440] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2442 &  data$isco_mainjob <= 2443] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2445] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2451] <- 13
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2460] <- 13
```

Socio-cultural semi-professionals (14)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2230] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2330 &  data$isco_mainjob <= 2332] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 2444] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2446 &  data$isco_mainjob <= 2450] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 2452 &  data$isco_mainjob <= 2455] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 3200] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3220 &  data$isco_mainjob <= 3224] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 3226] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3229 &  data$isco_mainjob <= 3340] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3460 &  data$isco_mainjob <= 3472] <- 14
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 3480] <- 14

```


Skilled service (15)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 3225] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3227 &  data$isco_mainjob <= 3228] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 3473 &  data$isco_mainjob <= 3475] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5000 &  data$isco_mainjob <= 5113] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 5122] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5131 &  data$isco_mainjob <= 5132] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5140 &  data$isco_mainjob <= 5141] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 5143] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5160 &  data$isco_mainjob <= 5220] <- 15
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 8323] <- 15
```


Low-skilled service (16)
```{r}
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5120 &  data$isco_mainjob <= 5121] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5123 &  data$isco_mainjob <= 5130] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 5133 &  data$isco_mainjob <= 5139] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 5142] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 5149] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 5230] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob == 8322] <- 16
data$class16_r1[data$selfem_mainjob == 1 & data$isco_mainjob >= 9100 &  data$isco_mainjob <= 9152] <- 16
```


Label categories
```{r}
data$class16_r1[data$class16_r1 == -9] <- NA
val_labels(data$class16_r1) <- c("Large employers" = 1,
                             "Self-employed professionals" = 2,
                             "Small business owners with employees" = 3,
                             "Small business owners without employees" = 4,
                             "Technical experts" = 5,
                             "Technicians" = 6,
                             "Skilled manual" = 7,
                             "Low-skilled manual" = 8,
                             "Higher-grade managers and administrators" = 9,
                             "Lower-grade managers and administrators" = 10,
                             "Skilled clerks" = 11,
                             "Unskilled clerks" = 12,
                             "Socio-cultural professionals" = 13,
                             "Socio-cultural semi-professionals" = 14,
                             "Skilled service" = 15,
                             "Low-skilled service" = 16)
var_label(data$class16_r1) <- "Respondent's Oesch class position - 16 classes"

table(data$class16_r1)
```

Collapse into 8-categories
```{r}
data$class8_r1 <- NA
data$class8_r1[data$class16_r1 <= 2] <- 1
data$class8_r1[data$class16_r1 == 3 | data$class16_r1 == 4] <- 2
data$class8_r1[data$class16_r1 == 5 | data$class16_r1 == 6] <- 3
data$class8_r1[data$class16_r1 == 7 | data$class16_r1 == 8] <- 4
data$class8_r1[data$class16_r1 == 9 | data$class16_r1 == 10] <- 5
data$class8_r1[data$class16_r1 == 11 | data$class16_r1 == 12] <- 6
data$class8_r1[data$class16_r1 == 13 | data$class16_r1 == 14] <- 7
data$class8_r1[data$class16_r1 == 15 | data$class16_r1 == 16] <- 8
val_labels(data$class8_r1) <- c("Self-employed professionals and large employers" = 1,
                            "Small business owners" = 2,
                            "Technical (semi-)professionals" = 3,
                            "Production workers" = 4,
                            "(Associate) managers" = 5,
                            "Clerks" = 6,
                            "Socio-cultural (semi-)professionals" = 7,
                            "Service workers" = 8)
var_label(data$class8_r1) <- "Respondent's Oesch class position - 8 classes"
table(data$class8_r1)
```

Collapse into 5-categories
```{r}
data$class5_r1 <- NA
data$class5_r1[data$class16_r1 <= 2 | data$class16_r1 == 5 | data$class16_r1 == 9 | data$class16_r1 == 13] <- 1
data$class5_r1[data$class16_r1 == 6 | data$class16_r1 == 10 | data$class16_r1 == 14] <- 2
data$class5_r1[data$class16_r1 == 3 | data$class16_r1 == 4] <- 3
data$class5_r1[data$class16_r1 == 7 | data$class16_r1 == 11 | data$class16_r1 == 15] <- 4
data$class5_r1[data$class16_r1 == 8 | data$class16_r1 == 12 | data$class16_r1 == 16] <- 5
val_labels(data$class5_r1) <- c("Higher-grade service class" = 1,
                            "Lower-grade service class" = 2,
                            "Small business owners" = 3,
                            "Skilled workers" = 4,
                            "Unskilled workers" = 5)
var_label(data$class5_r1) <- "Respondent's Oesch class position - 5 classes"

```

### Waves 6-10
We repeat the same process for waves 6-10. Recode the variable _isco08_ (isco08 4-digit):

```{r}
data$isco_mainjob <- data$isco08
data$isco_mainjob[is.na(data$isco_mainjob)] <- -9
var_label(data$isco_mainjob) <- "Current occupation of respondent - isco88 4-digit"
```


Recode employment status for respondents. First, we create the variable _emplrel_r2_
```{r}
data$emplrel_r2 <- data$emplrel
data$emplrel_r2[is.na(data$emplrel_r2)] <- 9
val_label(data$emplrel_r2, 9) <- "Missing"
table(data$emplrel_r2)

```

Second, we create the variable _emplno_r2_
```{r}
data$emplno_r2 <- data$emplno
data$emplno_r2[is.na(data$emplno_r2)] <- 0
data$emplno_r2[data$emplno_r2 >= 1 & data$emplno_r2 <= 9] <- 1
data$emplno_r2[data$emplno_r2 >= 10 & data$emplno_r2 <= 66665] <- 2
val_labels(data$emplno_r2) <- c("0 employees" = 0,
                           "1-9 employees" = 1,
                           "10+ employees" = 2)
table(data$emplno_r2)

```

Finally, by combining the last two, we create the variable _selfem_mainjob_
```{r}
data$selfem_mainjob <- NA
data$selfem_mainjob[data$emplrel_r2 == 1 | data$emplrel_r2 == 9] <- 1
data$selfem_mainjob[data$emplrel_r2 == 2 & data$emplno_r2 == 0] <- 2
data$selfem_mainjob[data$emplrel_r2 == 3] <- 2
data$selfem_mainjob[data$emplrel_r2 == 2 & data$emplno_r2 == 1] <- 3
data$selfem_mainjob[data$emplrel_r2 == 2 & data$emplno_r2 == 2] <- 4
val_labels(data$selfem_mainjob) <- c("Not self-employed" = 1,
                                  "Self-empl without employees" = 2,
                                  "Self-empl with 1-9 employees" = 3,
                                  "Self-empl with 10 or more" = 4)
var_label(data$selfem_mainjob) <- "Employment status for respondants"
table(data$selfem_mainjob)
```


Create Oesch class schema for respondents

We start creating the variable _class16_r2_
```{r}
data$class16_r2 <- -9

```

Large employers (1)
```{r}
data$class16_r2[data$selfem_mainjob == 4] <- 1

```

Self-employed professionals (2)

```{r}
data$class16_r2[(data$selfem_mainjob == 2 | data$selfem_mainjob == 3) & data$isco_mainjob >= 2000 & data$isco_mainjob <= 2229] <- 2
data$class16_r2[(data$selfem_mainjob == 2 | data$selfem_mainjob == 3) & data$isco_mainjob >= 2300 & data$isco_mainjob <= 2470] <- 2
```


Small business owners with employees (3)
```{r}
data$class16_r2[data$selfem_mainjob == 3 & data$isco_mainjob >= 1000 & data$isco_mainjob <= 1999] <- 3
data$class16_r2[data$selfem_mainjob == 3 & data$isco_mainjob >= 3000 & data$isco_mainjob <= 9333] <- 3
data$class16_r2[data$selfem_mainjob == 3 & data$isco_mainjob == 2230] <- 3

```

Small business owners without employees (4)

```{r}
data$class16_r2[data$selfem_mainjob == 2 & data$isco_mainjob >= 1000 & data$isco_mainjob <= 1999] <- 4
data$class16_r2[data$selfem_mainjob == 2 & data$isco_mainjob >= 3000 & data$isco_mainjob <= 9333] <- 4
data$class16_r2[data$selfem_mainjob == 2 & data$isco_mainjob == 2230] <- 4
```

Technical experts (5)

```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2100 &  data$isco_mainjob <= 2213] <- 5
```

Technicians (6)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3100 &  data$isco_mainjob <= 3152] <- 6
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3210 &  data$isco_mainjob <= 3213] <- 6
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 3434] <- 6
```


Skilled manual (7)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 6000 &  data$isco_mainjob <= 7442] <- 7
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 8310 &  data$isco_mainjob <= 8312] <- 7
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 8324 &  data$isco_mainjob <= 8330] <- 7
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 8332 &  data$isco_mainjob <= 8340] <- 7
```


Low-skilled manual (8)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 8000 &  data$isco_mainjob <= 8300] <- 8
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 8320 &  data$isco_mainjob <= 8321] <- 8
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 8331] <- 8
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 9153 &  data$isco_mainjob <= 9333] <- 8
```


Higher-grade managers and administrators (9)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 1000 &  data$isco_mainjob <= 1239] <- 9
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2400 &  data$isco_mainjob <= 2429] <- 9
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2441] <- 9
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2470] <- 9
```


Lower-grade managers and administrators (10)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 1300 &  data$isco_mainjob <= 1319] <- 10
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3400 &  data$isco_mainjob <= 3433] <- 10
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3440 &  data$isco_mainjob <= 3450] <- 10
```


Skilled clerks (11)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 4000 &  data$isco_mainjob <= 4112] <- 11
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 4114 &  data$isco_mainjob <= 4210] <- 11
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 4212 &  data$isco_mainjob <= 4222] <- 11
```

Unskilled clerks (12)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 4113] <- 12
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 4211] <- 12
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 4223] <- 12
```

Socio-cultural professionals (13)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2220 &  data$isco_mainjob <= 2229] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2300 &  data$isco_mainjob <= 2320] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2340 &  data$isco_mainjob <= 2359] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2430 &  data$isco_mainjob <= 2440] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2442 &  data$isco_mainjob <= 2443] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2445] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2451] <- 13
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2460] <- 13
```

Socio-cultural semi-professionals (14)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2230] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2330 &  data$isco_mainjob <= 2332] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 2444] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2446 &  data$isco_mainjob <= 2450] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 2452 &  data$isco_mainjob <= 2455] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 3200] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3220 &  data$isco_mainjob <= 3224] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 3226] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3229 &  data$isco_mainjob <= 3340] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3460 &  data$isco_mainjob <= 3472] <- 14
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 3480] <- 14

```


Skilled service (15)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 3225] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3227 &  data$isco_mainjob <= 3228] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 3473 &  data$isco_mainjob <= 3475] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5000 &  data$isco_mainjob <= 5113] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 5122] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5131 &  data$isco_mainjob <= 5132] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5140 &  data$isco_mainjob <= 5141] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 5143] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5160 &  data$isco_mainjob <= 5220] <- 15
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 8323] <- 15
```


Low-skilled service (16)
```{r}
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5120 &  data$isco_mainjob <= 5121] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5123 &  data$isco_mainjob <= 5130] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 5133 &  data$isco_mainjob <= 5139] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 5142] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 5149] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 5230] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob == 8322] <- 16
data$class16_r2[data$selfem_mainjob == 1 & data$isco_mainjob >= 9100 &  data$isco_mainjob <= 9152] <- 16
```


Label categories
```{r}
data$class16_r2[data$class16_r2 == -9] <- NA
val_labels(data$class16_r2) <- c("Large employers" = 1,
                             "Self-employed professionals" = 2,
                             "Small business owners with employees" = 3,
                             "Small business owners without employees" = 4,
                             "Technical experts" = 5,
                             "Technicians" = 6,
                             "Skilled manual" = 7,
                             "Low-skilled manual" = 8,
                             "Higher-grade managers and administrators" = 9,
                             "Lower-grade managers and administrators" = 10,
                             "Skilled clerks" = 11,
                             "Unskilled clerks" = 12,
                             "Socio-cultural professionals" = 13,
                             "Socio-cultural semi-professionals" = 14,
                             "Skilled service" = 15,
                             "Low-skilled service" = 16)
var_label(data$class16_r2) <- "Respondent's Oesch class position - 16 classes"

table(data$class16_r2)
```

Collapse into 8-categories
```{r}
data$class8_r2 <- NA
data$class8_r2[data$class16_r2 <= 2] <- 1
data$class8_r2[data$class16_r2 == 3 | data$class16_r2 == 4] <- 2
data$class8_r2[data$class16_r2 == 5 | data$class16_r2 == 6] <- 3
data$class8_r2[data$class16_r2 == 7 | data$class16_r2 == 8] <- 4
data$class8_r2[data$class16_r2 == 9 | data$class16_r2 == 10] <- 5
data$class8_r2[data$class16_r2 == 11 | data$class16_r2 == 12] <- 6
data$class8_r2[data$class16_r2 == 13 | data$class16_r2 == 14] <- 7
data$class8_r2[data$class16_r2 == 15 | data$class16_r2 == 16] <- 8
val_labels(data$class8_r2) <- c("Self-employed professionals and large employers" = 1,
                            "Small business owners" = 2,
                            "Technical (semi-)professionals" = 3,
                            "Production workers" = 4,
                            "(Associate) managers" = 5,
                            "Clerks" = 6,
                            "Socio-cultural (semi-)professionals" = 7,
                            "Service workers" = 8)
var_label(data$class8_r2) <- "Respondent's Oesch class position - 8 classes"
table(data$class8_r2)
```

Collapse into 5-categories
```{r}
data$class5_r2 <- NA
data$class5_r2[data$class16_r2 <= 2 | data$class16_r2 == 5 | data$class16_r2 == 9 | data$class16_r2 == 13] <- 1
data$class5_r2[data$class16_r2 == 6 | data$class16_r2 == 10 | data$class16_r2 == 14] <- 2
data$class5_r2[data$class16_r2 == 3 | data$class16_r2 == 4] <- 3
data$class5_r2[data$class16_r2 == 7 | data$class16_r2 == 11 | data$class16_r2 == 15] <- 4
data$class5_r2[data$class16_r2 == 8 | data$class16_r2 == 12 | data$class16_r2 == 16] <- 5
val_labels(data$class5_r2) <- c("Higher-grade service class" = 1,
                            "Lower-grade service class" = 2,
                            "Small business owners" = 3,
                            "Skilled workers" = 4,
                            "Unskilled workers" = 5)
var_label(data$class5_r2) <- "Respondent's Oesch class position - 5 classes"

```


### All waves
Create the variable _class5_r_ gathering both  _class5_r2_ and  _class5_r2_

```{r}
data$class5_r[data$essround<=10]<-data$class5_r2
data$class5_r[data$essround<6]<-data$class5_r1

# As factor 
data$class5_r <- factor(data$class5_r, 
                        levels=c(1,2,3,4,5), 
                        labels=c("Higher-grade service class",
                            "Lower-grade service class",
                            "Small business owners",
                            "Skilled workers",
                            "Unskilled workers"))
with(data, table(essround, class5_r))

```


Finally, we can extract R code from the R Markdown document using the `purl()` function from the `knitr` package. This function takes an R Markdown file as input and generates a pure R script containing only the code chunks from the document. Here's how you can do it:

```{r}
options(knitr.duplicate.label = 'allow')
purl("ESS_es_sociodemographics.Rmd", output = "extracted_data_wrangling_code.R")
```

By extracting the code into a standalone R script, you can easily reuse and source the code in other projects, R scripts, or R Markdown documents. This is especially useful for wrangling tasks, where you might develop a series of data cleaning, transformation, and analysis steps that are applicable across multiple datasets or projects.


