---
title: "Using Socio-Demographic Data in Practice"
author: "Marga Torre"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
---

<style>
body {
text-align: justify}
</style> 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, message=FALSE, warning=FALSE, knitr.purl.inline = TRUE)
```

```{r include=FALSE}
library(haven)
library(tidyverse)
library(knitr)
library(gmodels)
library(ggplot2)
library(openxlsx)
library(skimr)
library(ggplot2)
library(janitor)
library(readr)
library(dplyr)
library(labelled)
library(essurvey)
```

## European Social Survey
The European Social Survey (ESS) is an academically driven cross-national survey that has been conducted across Europe since its establishment in 2001. Every two years, face-to-face interviews are conducted with newly selected, cross-sectional samples. (Taken from the [ESS website](http://www.europeansocialsurvey.org/about-ess))


## Read data
```{r}
data_full <- read_csv("./ESS_es.csv")
dim(data_full)
```

You can use the [`essurvey` package](https://cran.r-project.org/web/packages/essurvey/index.html) to retrieve the data from the website. However, since it is not working properly at the moment while I'm preparing for the class, I have downloaded and imported the data in the conventional way.


### Select variables
ESS round + ID + socio-demographic profile, including: gender, age of the respondent, years of education, level of education, occupation, household's total net income and citizenship. Please be aware that __some variable names may change from one round to another__.

```{r}
data<- data_full |> select(essround, idno, gndr, agea, eduyrs, eisced, iscoco, isco08, hinctnt, hinctnta)
```

## Data cleaning and data wrangling 
### Gender

Generate a new variable named `men` and assign it the following values. Label the categories of the variable  and check the results using a table.

-   1=man
-   0=woman
-   other=NA

```{r}
data$men = if_else(data$gndr == 1, "men", 
                           if_else(data$gndr == 2, "women", NA))
table(data$men)
```

### Age

Generate a continuous variable named `age`, with the maximum value being 103 years:

```{r}
data = data |> mutate(age = replace(agea, agea>=103, NA))

summary(data$age)
```

Using the continuous variable `age`, generate a new variable named `age_rec` with the following categories (levels). Don't forget to label the categories of the variable and check the results using a table.

1.  "<20"
2.  "20-29"
3.  "30-44"
4.  "45-64"
5.  ">=65" 

```{r}
data <- data |> 
  mutate(age_rec = cut(agea, breaks = c(0, 19, 29, 44, 64, 103),
                       labels = c("<20", "20-29", "30-44", "45-64", ">=65")))

with(data, (table(essround, age_rec))) # it indicates the variables name in the table
table(data$essround, data$age_rec) # it does not have variables name in the table
```

### Education
Generate a variable named `years_schooling` defined as a continuous variable. 

```{r}
data <- data |> mutate(years_schooling = if_else(data$eduyrs <77, eduyrs, NA))
with(data, table(essround, years_schooling))
```

Based on the _EISCED_ variable, generate a variable named _educ_level_ with the below categories. Ensure that the categories of the new variable are labeled  and check the results using a table.

1.  Low education
2.  Medium education
3.  High education

```{r}
data = data |> mutate(educ_level = 
                        case_when(
                          eisced %in% c(1,2) ~ "1",
                          eisced %in% c(3,4) ~ "2",
                          eisced %in% c(5,6,7) ~ "3",
                          eisced %in% c(55,77,88,99) ~ NA,
                          TRUE ~ NA
                        )) |> 
  mutate(educ_level = fct_recode(educ_level,
                                   "Low education" = "1",
                                   "Medium education" = "2",
                                   "High eudcation" = "3"))



table(data$educ_level)
```


### Household's total net income
Generate a scale variable named hincome displaying income deciles. Note that the variable name changes from wave 3 to wave 4.

```{r}
data <- data |> mutate(hincome = 
                         case_when(
                           essround <= 3 ~ hinctnt,
                           essround >=4 ~ hinctnta,
                           TRUE ~ NA
                         ))

data <- data |> relocate(hincome, .before = hinctnt)
with(data, table(essround, hincome))

# ntile(head(data$hinctnt), 10)
# ntile(tail(data$hinctnta), 10)
```


### Occupational categories
Using the variables _iscoco_ (waves <=5) and _isco08_ (waves >5), generate a new variable named `occup_cat` with the following levels:

1.    Managers
2.    Professionals
3.    Technicians
4.    Clerical
5.    Service
6.    Agricultural
7.    Blue-collar


```{r}
data <- data |> mutate(occup_cat = 
                         case_when(
                           essround <=5 ~ iscoco,
                           essround >5 ~ isco08,
                           TRUE ~ NA
                         ))


data <- data |> 
  mutate(
    occup_cat = replace(occup_cat, which(occup_cat > 10000), NA),
    occup_cat = cut(
      occup_cat, breaks = c(0, 1999, 2999, 3999, 4999, 5999, 6999, 9999),
      labels = c(
        "managers",
        "professionals",
        "technicians",
        "clerical",
        "service",
        "agricultural",
        "blue-collar"
      )
    )
  )

table(data$occup_cat)
```


## OPTIONAL: Social class - Daniel Oesch 

Replicate Daniel Oesch's files __to construct the 5-class schemes__ for the interviewee's class position. Please note that starting from wave 6, there has been a change in the occupation classification from `iscoco` to `isco08`. Use [the scripts provided by Daniel Oesch](https://people.unil.ch/danieloesch/scripts/)

References:

-   Oesch, D. (2006a) "Coming to grips with a changing class structure" _International Sociology_ 21 (2): 263-288.
-   Oesch, D. (2006b) "Redrawing the Class Map. Stratification and Institutions in Britain, Germany, Sweden and Switzerland", Basingstoke: Palgrave Macmillan.

```{r}
data2 <- data_full |> select (idno, essround, emplrel, emplno) 
data <- left_join(data, data2, by=c("essround", "idno"))

data$isco_mainjob <- data$iscoco
data$isco_mainjob[is.na(data$isco_mainjob)] <- -9
var_label(data$isco_mainjob) <- "Current occupation of respondent - isco88 4-digit"
```



#### 5-Class schema constructed

1.    Higher-grade service class
2.    Lower-grade service class
3.    Small business owners
4.    Skilled workers
5.    Unskilled workers

