---
title: "Exercise: using Paradata"
author: "Marga Torre"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged

---

```{=html}
<style>
body {
text-align: justify}
</style>
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, message=FALSE, warning=FALSE, knitr.purl.inline = TRUE )
```

```{r include=F}


```

## Intro
Individuals' ideology is a crucial element in predicting voting patterns, yet it often has the highest non-response rate, complicating electoral forecasts. 

Using both the __ESS core data__ and the __Interviewer Data File__, __Round 10__ (2020-2022) for __all countries,__ examine the following questions:

-   Is the non-response rate higher in some contexts than in others? 
-   What profile of interviewee is more likely to not reveal their ideology? 
-   How does the interviewer's profile influence non-response, if at all?

## Read data
Load both the core data and the interviewer's data file (Round 10) for all countries. 
```{r}
data <- read.csv("./ESS10.csv")
paradata <- read.csv("./ESS10_INT.csv")
```

Merge the two datasets by country and identification number
```{r}
library(dplyr)
library(stringr)
data <- left_join(data, paradata, by = c("idno", "cntry")) |> 
  rename_with(~str_replace(., "\\.y$", ""), .cols = ends_with(".y")) |> 
  select(-ends_with(".x"))
# data = data |> select(-contains("x"))
head(data) 
```


## Diferences by country
Examine the distribution of missing values (NA) in the variable ideology by country. 
```{r}
data = data |> 
  mutate(ideology_na = ifelse(lrscale > 10, 1, 0))

prob_table = prob.table(table(data$cntry, data$ideology_na), margin=1)*100
```

## The effect of individual's profile
Prepare the gender and age of the individual for the analysis
```{r}
# gender (man/woman)
data <- data |>
  mutate(gender = factor(ifelse(gndr == 1, "man", ifelse(gndr == 2, "woman", NA))))
table(data$gender)

# age
data <- data |>
  mutate(age = ifelse(agea == 999, NA, agea))
summary(data$age)
```

Now, examine the probability of non-response based on the gender and age of the interviewer. Discuss the results.

```{r}
m0 = glm(ideology_na ~ gender + age + factor(cntry), data, family = "binomial")
summary(m0)
```

Examine if age has a non-linear effect.  Discuss the results

```{r}
m0b = glm(ideology_na ~ gender + age + I(age^2) + factor(cntry), data, family = binomial(link = "logit"))
summary(m0b)
```

## The effect of interviewer's profile
Now, utilize paradata to investigate the impact of the interviewer's profile on the likelihood of non-response. Specifically, analyze the effects of the interviewer's gender, age, and interviewer's overall experience conducting the survey

```{r}
# gender of the interviewer (man/woman)
data <- data |>
  mutate(intgender = factor(ifelse(intgndr == 1, "man", ifelse(intgndr == 2, "woman", NA))))
table(data$intgender)

# age of the interviewer
data <- data |>
  mutate(intage = ifelse(intagea == 999, NA, intagea))
summary(data$intage)

# overall experience conducting the interview
table(data$intgender, data$intage)
data = data |> mutate(intexperience = ifelse(vdovexin >10, NA, vdovexin))
summary(data$intexperience)
```


Run the regression and discuss the results
```{r}
m1<-glm(ideology_na ~ gender + age  + intgender + intage + + intexperience + factor(cntry), data, family=binomial(link = "logit"))
summary(m1)
```

Note the impact of allowing age to have a non-linear effect. Would you say there is a maximum or minimum in the effect of age? Explain your answer. 
```{r}
m1b<-glm(ideology_na ~ gender + age + I(age^2) + intgender + intage + I(intage^2) + intexperience + factor(cntry), data, family=binomial(link = "logit"))
summary(m1b)
```

Finally enhance the regression analysis by introducing a new variable that quantifies the gender match (or mismatch) between the interviewer and the interviewee.

1.    mixed
2.    female-female
3.    male-male

```{r}
data <- data |>
  filter(gndr <= 9, intgndr <= 9) |>
  mutate(gender_combination = factor(case_when(
    gndr == 1 & intgndr == 1 ~ "male-male",    # male-male
    gndr == 2 & intgndr == 2 ~ "female-female", # female-female
    gndr == 1 & intgndr == 2 ~ "mixed",   # male-female
    gndr == 2 & intgndr == 1 ~ "mixed"    # female-male
  )))

# Table of gender_combination
table(data$gender_combination)
```

Run the regression and discuss the results.
```{r}
m1c<-glm(ideology_na ~ age + I(age^2) + gender_combination + intage + I(intage^2) + intexperience + factor(cntry), data, family=binomial(link = "logit"))
summary(m1c)
```


## Extracting tables from R statistical output

```{r}
# jtools
install.packages("stargazer")
library(stargazer)

stargazer(m1b, m1c,  
          type="text",
          dep.var.labels=c("Probability of non-response in ideology"),
          keep = c(
            "(Intercept)", 
            "age", 
            "I(age^2)", 
            "gender_combinationmale-male", 
            "gender_combinationmixed", 
            "intage", 
            "I(intage^2)", 
            "intexperience"),
          covariate.labels = c(
            "Age of the interviewee",
            "Age of the interviewee (squared)",
            "Pair: male-male (rc: female-famale)",
            "Pair: mixed (rc: female-famale)",
            "Age of the interviewer", 
            "Age of the interviewer (squared)", 
            "Overall experience of the interviewer"),        
          out="models.txt")
```







